---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: TRUE
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
Read the csv file and convert the column date from class character to date

```{r readdata, cache=FALSE}
library(tidyverse)
library(simputation)
library(knitr)
data<-read.csv("activity.csv" , sep = ",")
data$date<-as.Date(data$date, format = "%Y-%m-%d")  
```
## What is mean total number of steps taken per day?

```{r meanmedian, cache=TRUE}
stepspday<-data%>%group_by(date)%>%summarise(mean=mean(steps,na.rm = T), median=median(steps, na.rm = T),sum=sum(steps,na.rm = T))
meansteps<-mean(stepspday$sum, na.rm = T)
meansteps
mediansteps<-median(stepspday$sum, na.rm = T)
mediansteps
```

```{r stepsperday, include=TRUE}
with(stepspday, hist(sum) )
```

## What is the average daily activity pattern?

```{r pattern}

timeseries<-data%>%group_by(interval)%>%summarize(mean_steps=mean(steps, na.rm = T))
timeseries%>%ggplot(aes(interval, mean_steps)) + geom_line()
```

The 5-minute interval that, on average, contains the maximum number of steps is:
```{r maxint, cache=TRUE}
maxsteps<-which.max(timeseries$mean_steps)
timeseries[maxsteps,1]

```

## Imputing missing values
##Code to describe and show a strategy for imputing missing data


Downloaded the package simputation and found the function impute_proxy
Method: imputation by a group mean, calculates the mean of each group (interval ) and replaces the NA within each group with the mean
```{r imputemissingdata, cache=TRUE}
summary(data)
datanew<-impute_proxy(dat = data,formula = steps~mean(steps,na.rm = T)| interval)
summary(datanew)

```

##Histogram of the total number of steps taken each day after missing values are imputed
```{r missing}
data1<-datanew%>%group_by(date)%>%summarise(mean=mean(steps,na.rm = T), median=median(steps, na.rm = T),sum=sum(steps,na.rm = T))
with(data1, hist(sum) )
meanstepsnona<-mean(data1$sum)
meanstepsnona
medianstepsnona<-median(data1$sum)
medianstepsnona

```


## Are there differences in activity patterns between weekdays and weekends?
```{r compare}
data2<-datanew%>%mutate (day = weekdays(date) )
data2<-data2%>%mutate(dayw=ifelse(day %in% c("sábado", "domingo"), "weekend", "weekday"))

data3<-data2%>%group_by(interval)%>%summarise(interval,mean_steps_day=mean(steps),dayw)
data3%>%ggplot(aes(interval, mean_steps_day)) +geom_line()+facet_grid(dayw~.)



```
